\normalfont{A statistical analysis on factors that contributed/slowed down the spread of COVID-19}

### 2020-09-08 Manuscript/Manuscript.tex
PUB

\documentclass[
12pt, % Main document font size
a4paper, % Paper type, use 'letterpaper' for US Letter paper
oneside, % One page layout (no page indentation)
%twoside, % Two page layout (page indentation for binding and different headers)
headinclude,footinclude, % Extra spacing for the header and footer
BCOR5mm, % Binding correction
]{scrartcl}
\usepackage[english]{babel}
\usepackage{url}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{float}
\usepackage{epigraph}
\usepackage{mathcomp}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{url}

@@

\newpage

\epigraph{ "Text" }

@@

\listoffigures % Print the list of figures

@@

%----------------------------------------------------------------------------------------
%	AUTHOR AFFILIATIONS
%----------------------------------------------------------------------------------------

%\let\thefootnote\relax\footnotetext{* \textit{Department of Biology, University of Examples, London, United Kingdom}}

%\let\thefootnote\relax\footnotetext{\textsuperscript{1} \textit{Department of Chemistry, University of Examples, London, United Kingdom}}

@@

\section{Theoretical Background} \label{Theoretical Background}
As pointed out by Mackay \cite{mackay2003information} the basic idea of unsupervised learning is to mime human behaviour is to find regularities in data and group them. Among the different techniques of unsupervised learning \cite{james2013introduction} here we considered the PCA: in this section the basic ideas of this tool will be focused following the approach proposed by James et al. \cite{james2013introduction}. The aim of PCA is to plot $n$ observations with $p$ features in only one 2D plot,with the least possible loss of information, instead of $\binom{p}{2}$ 2D plots. To do so the PCA chooses, among the possible axis formed by the linear normalized combination of the features, the two associated with the largest variance, given that these two axis are orthogonal. Thus, supposing that the means of the features are null, the problem is to find the coefficients (usually called loadings) $\phi_{ij}$ that solves the following optimization problem:

@@

with the constraint

@@

The loadings define the principal components (that have to be orthogonal) :

@@

If the all the principal components are taken in to account no information is lost, however the complexity of it due to high dimensionality is the same of the starting one; however if we sacrifice the p-2 component, that have a lower variance with respect to the first two, the complexity is dramatically reduced to a 2D plot. In this case the toll paid is the lost of the information contained in the $p-2$ component. From a computational point of view this task can be performed with the standard techniques used for solving a eigenvalue problem. Once the two principal component are found, the data can plotted into the new coordinate set; the advantage of this new representation lies on the fact that if the first two principal component loading vectors are plotted a bird's-eye view of the statistical correlation between the features is obtained: basically the cosine of the angle between the loadings approximates the statistical correlation, while the position of the data with respect to the loadings will plot its feature. In this way an intuitive representation of the data, their features and the correlation among them is provided. Finally this can be accompanied by a plot of the correlation matrix as it is done in this work.

@@

\subsection{Provinces}

@@

\begin{figure}[h]
\begin{center}
\includegraphics[scale=1]{Pic/CorrMatrix_WORLD.pdf}
\caption{The Correlation matrix for the a set of countries accompanied by its heat-map as generated by the following dataset:  }
\label{World_CorrMatrix}
\end{center}
\end{figure}

@@

%------------------------------------------------

>>>>

\documentclass[
12pt, % Main document font size
a4paper, % Paper type, use 'letterpaper' for US Letter paper
oneside, % One page layout (no page indentation)
%twoside, % Two page layout (page indentation for binding and different headers)
headinclude,footinclude, % Extra spacing for the header and footer
BCOR5mm, % Binding correction
]{scrartcl}
\usepackage[english]{babel}
\usepackage{url}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{float}
\usepackage{epigraph}
\usepackage{mathcomp}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{url}
\usepackage{sectsty}
\usepackage[dvipsnames,table,xcdraw,svgnames]{xcolor}
\usepackage{listings}
\lstset{language=R,
    basicstyle=\small\ttfamily,
    stringstyle=\color{DarkGreen},
    otherkeywords={0,1,2,3,4,5,6,7,8,9},
    morekeywords={TRUE,FALSE},
    deletekeywords={data,frame,length,as,character},
    keywordstyle=\color{blue},
    commentstyle=\color{DarkGreen},
}

@@

\newpage
\setlength\epigraphwidth{.7\textwidth}
\epigraph{ "Haec ratio quondam morborum et mortifer aestus \\
finibus in Cecropis funestos reddidit agros \\
vastavitque vias, exhausit civibus urbem. \\
nam penitus veniens Aegypti finibus ortus, \\
aeëra permensus multum camposque natantis, \\
incubuit tandem populo Pandionis omni." \\
Lucretius, De Rerum Natura VI \\
\vspace{5mm}
Twas such a manner of disease, 'twas such \\
Mortal miasma in Cecropian lands \\
Whilom reduced the plains to dead men's bones,\\
Unpeopled the highways, drained of citizens \\
The Athenian town. For coming from afar,\\
Rising in lands of Aegypt, traversing \\
Reaches of air and floating fields of foam, \\
At last on all Pandion's folk it swooped \\
(As translated by William Ellery Leonard)
}

@@

%\listoffigures % Print the list of figures

@@

@@

\section{Theoretical Background} \label{Theoretical Background}
As pointed out by Mackay \cite{mackay2003information} the basic idea of unsupervised learning is to mime human behaviour is to find regularities in data and group them. Among the different techniques of unsupervised learning \cite{james2013introduction} here we considered the PCA: in this section the basic ideas of this tool will be focused following the approach proposed by James et al. \cite{james2013introduction}. The aim of PCA is to plot $n$ observations with $p$ features in only one 2D plot,with the least possible loss of information, instead of $\binom{p}{2}$ 2D plots. To do so the PCA chooses, among the possible axis formed by the linear normalized combination of the features, the two associated with the largest variance, given that these two axis are orthogonal. Thus, supposing that the means of the features are null, the problem is to find the coefficients (usually called loadings) $\phi_{ij}$ that solves the following optimization problem  \cite{james2013introduction}:

@@

with the constraint  \cite{james2013introduction}

@@

The loadings define the principal components (that have to be orthogonal)  \cite{james2013introduction}:

@@

If the all the principal components are taken in to account no information is lost, however the complexity of it due to high dimensionality is the same of the starting one; however if we sacrifice the p-2 component, that have a lower variance with respect to the first two, the complexity is dramatically reduced to a 2D plot. In this case the toll paid is the lost of the information contained in the p-2 components. From a computational point of view this task can be performed with the standard techniques used for solving a eigenvalue problem. Once the two principal component are found, the data can plotted into the new coordinate set; the advantage of this new representation lies on the fact that if the first two principal component loading vectors are plotted a bird's-eye view of the statistical correlation between the features is obtained: basically the cosine of the angle between the loadings approximates the statistical correlation, while the position of the data with respect to the loadings will plot its feature. In this way an intuitive representation of the data, their features and the correlation among them is provided. Finally this can be accompanied by a plot of the correlation matrix as it is done in this work.

@@

\subsection{Provinces} \label{sub_prov}

@@

For this dataset it is possible to see from the inspection of Fig. \ref{World_CorrMatrix}, \ref{World_PCA_loadings} and \ref{World_FULL_PCA} that among the different features considered for the countries only the GDP seems to be statistically correlated with the cases and the deaths. This fact can be explained with the same argument discussed in the section \ref{sub_prov}. In principle a possible objection to this link can be that only the richest countries can invest money for the tests for the validation of a COVID19 case: however it can be noted from plots that the normalized health expenditure as well the UHC index are basically orthogonal with respect to the normalized COVID19 cases and deaths. Moreover it is rummy, on the basis of recent papers , to find that the COVID cases and deaths are orthogonal to the pollution: this fact may be ascribed to the fact that the a smaller cluster should be considered; however it is worth nothing that in the literature the a strong causal relation is far from being validated \cite{contini2020does}. Finally, from the inspection of Fig. \ref{World_variances}, it is possibile to point out that the in the 2D plots \ref{Regioni_PCA_loadings} and \ref{Regions_FULL_PCA} the percentage of variance reported is almost the 58 $ \% $ while the 42 $ \% $ is lost.

\begin{figure}[h]
\begin{center}
\includegraphics[scale=1]{Pic/CorrMatrix_WORLD.pdf}
\caption{The Correlation matrix for the a selected set of countries accompanied by its heat-map as generated by the following dataset: The PM 2.5 concentration ($\mu g/m^{3}$ 2016), the normalized traffic morality (2016), the normalized pollution mortality, the GDP pro-capita(2019), the normalized deaths for COVID19 (27/08/2020), the overall number of cases (27/08/2020), the density (2019), the normalised health expenditure (2017), and the Universal Health Coverage Index (2017)}
\label{World_CorrMatrix}
\end{center}
\end{figure}

@@

\section{Conclusions}

The analysis of the features of the different clusters pinpointed the following features:

\begin{itemize}
\item The normalized number of cases and deaths for COVID19 are positively correlated for all cluster with the mean income. As validated by the previous literature (see the subsection \ref{sub_prov}) , this relation is causal due to the fact a higher income increase the connectivity of the agents
\item The unemployment is negatively correlated with the number of cases and deaths: this effect is entangled with the mean income; thus for a deeper analysis on the causal relationship, an econometric research that disentangle the mean income and the employment is required. 
\item The spread of COVID19 does not seems to be negatively correlated with the quality and quantity of hospitals (LEA, UHC, Number of public structure), in reverse the number of general practitioner and their visit is negatively correlated with the number of cases and deaths. This can be due to the fact the a large number of normalized general practitioner are able to detect and isolate the infected agents, and this slows the spread of a disease. However a deeper analysis on this statical correlation is required in order support this argument. 
\item The public transport is positively correlated with the normalized new cases: the causal reason for this is that the agents during the travel increase their interactions with other agents thus the probability to be infected is higher
\end{itemize}

\section{R code}
\begin{lstlisting}
# Provinces
Dati_Covid <- data.frame
(read_excel("C:/Users/Marzio/Desktop/Progetti ML/Dati Covid_prov_FULL.xlsx"))

row.names(Dati_Covid)<- Dati_Covid[,2]

prov_full.pca<-prcomp(Dati_Covid[,5:11], center = TRUE,scale. = TRUE)

summary(prov_full.pca)

autoplot(prov_full.pca,loadings = TRUE,loadings.size= 4,
loadings.colour = 'blue',
loadings.label = TRUE, loadings.label.size = 4,
label=TRUE,label.vjust=1.5,
loadings.label.repel=T,label.repel=T)
+ggtitle('Totale') + theme_bw()
+ theme(plot.title = element_text(hjust = 0.5))

fviz_eig(prov_full.pca,barcolor = "red",
barfill = "red",geom = c("bar"),addlabels= TRUE )
+labs(title = "Variances - PCA",
x = "Principal Components", y = "% of variances")
+theme_bw()+theme(plot.title = element_text(hjust = 0.5))

fviz_pca_var(prov_full.pca,col.var = "contrib",
gradient.cols = c("red","orange","blue"),
repel = TRUE,col.circle = "black",arrowsize = 1,labelsize = 0.5,
jitter = list(what = "both", width = 1, height = 1) ) 
+theme_bw()+theme(plot.title = element_text(hjust = 0.5))

corrplot(cor(Dati_Covid[,5:11]))
palette = colorRampPalette(c("green", "blue", "red")) (20)
heatmap(x = cor(Dati_Covid[,5:11]), col = palette, 
symm = TRUE, margins = c(10, 10),
 main = 'Provinces',dist(Dati_Covid[,5:11],method = 'euclidean'))

# Regions
Dati_Covid_reg <- data.frame(read_excel
("C:/Users/Marzio/Desktop/Progetti ML/Dati Covid reg_FULL.xlsx"))

row.names(Dati_Covid_reg) <-Dati_Covid_reg[,1]

reg_full.pca<-prcomp(Dati_Covid_reg[,2:11], center = TRUE,scale. = TRUE)

autoplot(reg_full.pca,loadings = TRUE,
loadings.size= 4, loadings.colour = 'blue',
loadings.label = TRUE, loadings.label.size = 4,
label=FALSE,label.vjust=1.5,
loadings.label.repel=T,label.repel=T)
+ggtitle('Regions') +theme_bw()+
theme(plot.title = element_text(hjust = 0.5))

summary(reg_full.pca)

fviz_eig(reg_full.pca,barcolor = "red",
barfill = "red",geom = c("bar"),addlabels= TRUE )
+labs(title = "Variances - PCA",
x = "Principal Components", y = "% of variances")
+theme_bw()+theme(plot.title = element_text(hjust = 0.5))
fviz_pca_var(reg_full.pca,col.var = "contrib",
gradient.cols = c("red","orange","blue"),
repel = TRUE,col.circle = "black",arrowsize = 1,
labelsize = 0.5,jitter = list(what = "both", width = 1, height = 1) )
 +theme_bw()
+theme(plot.title = element_text(hjust = 0.5))

corrplot(cor(Dati_Covid_reg[,2:11]))
palette = colorRampPalette(c("green", "blue", "red")) (20)

heatmap(x = cor(Dati_Covid_reg[,2:11]),
 col = palette, symm = TRUE, margins = c(10, 10),
 main = 'Regions', dist(Dati_Covid_reg[,2:11],method = 'euclidean'))
dend.dat<-dist(Dati_Covid_reg[,2:11] )

#World
Covid_mond <- data.frame(
read_excel("C:/Users/Marzio/Desktop/Progetti ML/Covid_mond.xlsx"))

row.names(Covid_mond) <-Covid_mond[,1]

world.pca<-prcomp(Covid_mond[,6:14], center = TRUE,scale. = TRUE)

fviz_pca_var(world.pca,col.var = "contrib",
gradient.cols = c("red","orange","blue"),
repel = TRUE,col.circle = "black",arrowsize = 1,
labelsize = 0.5,jitter = list(what = "both", width = 1, height = 1) ) 
+theme_bw()+theme(plot.title = element_text(hjust = 0.5)) 
fviz_eig(world.pca,barcolor = "red",
barfill = "red",geom = c("bar"),addlabels= TRUE )
+labs(title = "Variances - PCA",
x = "Principal Components", y = "% of variances")
+theme_bw()+theme(plot.title = element_text(hjust = 0.5))

autoplot(world.pca,loadings = TRUE,loadings.size= 4,
loadings.colour = 'blue',
loadings.label = TRUE,
loadings.label.size = 5,label=FALSE,label.vjust=1.5,
loadings.label.repel=T,label.repel=T)
+ggtitle('World')
+theme_bw()+
theme(plot.title = element_text(hjust = 0.5)) 
summary(world.pca)

corrplot(cor(Covid_mond[,6:14]))

palette = colorRampPalette(c("green", "blue", "red")) (20)

heatmap(x = cor(Covid_mond[,6:14]), 
col = palette, symm = TRUE, margins = c(10, 10),
main = 'World',dist(Covid_mond[,6:14],method = 'euclidean'))

\end{lstlisting}
%------------------------------------------------
